---
title: "PA1_template"
author: "Maureen McAndrew"
date: "4/9/2021"
output:
     html_document: default
     keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r chunk1}
unzip("C://Users/maure/Documents/R/Course 5 JHU Week 2 project/repdata_data_activity.zip")
reprodata <- read.csv("activity.csv",  header = TRUE, sep = ",")
```
change class of data variable
```{r chunk2}
library(lubridate)
reprodata$date <- as.Date(reprodata$date)
```
create a histogram of total steps per day
```{r hist1}
library(dplyr)
steps_day <- reprodata %>% group_by(date) %>% summarize(totalsteps = sum(steps))
steps_day_hist <- hist(steps_day$totalsteps, main = "Total Steps per Day")
```
Calculate mean and median number of steps per day
```{r chunk4}
a <- mean(steps_day$totalsteps, na.rm = TRUE)
b <- median(steps_day$totalsteps, na.rm = TRUE)
```
Remove na values from reprodata
```{r chunk5}
nonareprodata <- na.omit(reprodata)
```
Get mean number of steps taken averaged across all days 
```{r chunk6}
nonareprodata1 <- nonareprodata %>% group_by(date) %>% summarize(mean = mean(steps))
```
Create a time series plot of average steps per day
```{r plot2}
library(ggplot2)
plot2 <- ggplot(nonareprodata1, aes(date, mean))+
geom_line()+
geom_point()+
scale_x_date(date_labels = "%Y %b %d")+
labs(title = "Average Number of Steps Taken per Day", 
  x = "Date", 
  y = "Mean Steps per Day")
print(plot2)
```
Determine the 5 minute interval that gives the maximum number of steps
```{r chunk8}
library(dplyr)
max <- nonareprodata %>% group_by(interval, date) %>% summarize(max_steps = max(steps))
range (max$max_steps)
max %>% filter(max_steps == 806)
```
Create a histogram of total steps per day with imputed (used mean) missing 
values. Then calculate new median and mean of total steps for imputed dataset.
```{r hist2}
library(dplyr)
reprodata$steps [is.na(reprodata$steps)] <- mean(reprodata$steps, na.rm = TRUE)
steps_day_impute <- reprodata %>% group_by(date) %>% summarize(totalsteps = sum(steps))
hist_imputed <- hist(steps_day_impute$totalsteps, main = "Total Steps Per Day with Imputed Values", xlab = "Total Steps per day")
print(hist_imputed)
```
Get mean and median of total steps with imputed values
```{chunk10}
c <- mean(steps_day_impute$totalsteps)
d <- median(steps_day_impute$totalsteps)
c - a
d - b
# no mean difference between imputed values and removed na, but median difference is 1.188679
```
To create comparison panel plot between weekend and weekday values, first must separate dates into weekend and weekdays
```{r chunk11}
reprodata$date <- as.Date(reprodata$date)
library(dplyr)
reprodataw <- reprodata %>% mutate(weekday = weekdays(date))
reproweekday <- reprodataw %>% mutate(day = factor(ifelse(weekday == 
      c("Saturday", "Sunday"), "Weekend", "Weekday")))
```
Get mean number of steps by weekend or weekday according to time interval
```{r chunk12}
weekday <- reproweekday %>% group_by (interval, day) %>% summarize(mean_steps = mean(steps))
```
Create panel plot comparing time intervals for weekdays vs. weekends
```{r pplot}
library(ggplot2)
pplot <- ggplot(weekday, aes(interval, mean_steps, color = day))+
geom_line()+
facet_wrap(~day)+
ggtitle("Average Steps per 5 Minute Interval")+
xlab ("Time Intervals")+
ylab("Average Number of Steps")
print(pplot)
```
